---
title: "Homework 1"
author: Rachel Tao
output: github_document
---

## Problem 1

```{r load_packages}
library(tidyverse)
```

Create a data frame with specified elements:

```{r create_df}
prob_1_df <- 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("low", "low", "low", "mod","mod", "mod", "mod", "high", "high", "high"))
  )
```

Take the mean of each variable in the data frame:

```{r mean_var}
mean(pull(prob_1_df, samp))
mean(pull(prob_1_df, samp_gt_0))
mean(pull(prob_1_df, char_vec))
mean(pull(prob_1_df, factor_vec))
```

I can take the mean of numbers and logical but not character factor.

Convert all variables to numeric type:

```{r convert_numeric, results = FALSE, warning = FALSE}
as.numeric(pull(prob_1_df, samp))
as.numeric(pull(prob_1_df, samp_gt_0))
as.numeric(pull(prob_1_df, char_vec))
as.numeric(pull(prob_1_df, factor_vec))
```

When I apply as.numeric to the numeric, logical, and factor vectors, I get 
numbers, but when I apply as.numeric to the character vector, I get NA 10x. This
helps explain why I could not take the mean of the character vector - character 
does not convert to numeric.

Convert logical variable to different variable types (numeric, factor) multiply:

```{r convert_var_type}
#convert logical to numeric and multiply by numeric
as.numeric(pull(prob_1_df, samp_gt_0)) * pull(prob_1_df, samp)

#convert logical to factor and multiply by numeric
as.factor(pull(prob_1_df, samp_gt_0)) * pull(prob_1_df, samp)

#convert logical to factor, then convert the converted vector to numeric, multiply by numeric
as.numeric(as.factor(pull(prob_1_df, samp_gt_0))) * pull(prob_1_df, samp)
```

When I convert a logical vector to numeric, I can treat the converted vector as 
a numeric vector in a mathematical expression, and get numeric output. R treats factor 
vectors as non-numeric automatically, resulting in "NA" output when I convert a logical 
vector to factor and try to use it in a mathematical expression. However, it is 
possible to convert the factor to numeric and use this twice converted vector 
in a mathematical expression.

## Problem 2

Load penguins dataset:

```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

Describe penguins dataset:

```{r describe_penguins, echo=TRUE}

#view data and variables in dataset penguins
view(penguins)
head(penguins)

#rows in dataset penguins
nrow(penguins)
#columns in dataset penguins
ncol(penguins)

mean_flipper_length <- mean(pull(penguins, flipper_length_mm), na.rm = TRUE)
```

The penguins dataset has 344 rows and 8 columns. The variables are: 1) species, 2) island, 3) bill length (mm), 4) bill depth (mm), 5) flipper length (mm), 6) body mass (g), 7) sex, and 8) year. The mean flipper length is `r round(mean_flipper_length, digits=2)`.

Make a scatterplot of flipper length vs. bill length:

```{r scatter_flipper_bill}
prob_2_scatter <- ggplot(
  data = penguins, 
  mapping = aes(
    x = bill_length_mm, 
    y = flipper_length_mm, 
    color = species)) + 
  geom_point(na.rm = TRUE)

prob_2_scatter

#export plot to directory
ggsave("prob2_scatter.pdf", height = 4, width = 6)
```
